\documentclass{llncs}

\usepackage{amscd,amssymb}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{xspace}
%\usepackage[pdfusetitle,colorlinks,plainpages=false]{hyperref}
%\usepackage{biblatex}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{tikz-uml}
\usepackage[caption=false]{subfig}
\usepackage{tabularx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{units}
%\usepackage{amsthm}
\usepackage{algpseudocode}
\usepackage{algorithm}
%\usepackage[T1]{fontenc}
%\usepackage{lmodern}

\newenvironment{keywords}{
       \list{}{\advance\topsep by0.35cm\relax\small
       \leftmargin=0cm
       \labelwidth=0cm
       \listparindent=0cm
       \itemindent
       \listparindent
       \rightmargin\leftmargin
       }
  \item[\hskip\labelsep
                                     \bfseries Keywords:]}
     {\endlist}

\input{macros}

\begin{document}

\title{A Zero-Knowledge Proofs Compiler for Hardware-Software Co-Design}

\author{
Boran Car \and Josep Balasch \and Alfredo Rial \and Ingrid Verbauwhede \and Bart Preneel}
\institute{
IBBT and KU Leuven, ESAT-COSIC\\
\email{boran.car@student.kuleuven.be}\\
\email{\{firstname.lastname\}@esat.kuleuven.be}
}

\date{}
\maketitle

\input{abstract}

%\pagestyle{headings}
\section{Introduction}
\label{introduction}
\input{introduction}


\section{Related Work}
\label{relatedwork}
\input{relatedwork}


\section{Background}
In this section we review necessary background information. In the first part, we describe the targeted ZKPK $\Sigma$-protocols. In the second part, we introduce the concepts of syntax trees and parsers, and we finish by providing
definitions of data flow graph and control flow graph.
\label{preliminaries}
\input{preliminaries}


%\section{CACE ZKPK Compiler}
%\label{cacecompiler}

\section{Tools}
\label{sec:tools}
In this Section we introduce the different tools that conform the core
of our ZKPK framework. These include the CACE ZKPK compiler, the LLVM
compiler suite and the GEZEL hardware description language. A brief description
of the ANTLR parser generator is given in Appendix~\ref{sec:antlr}.
\input{cacecompiler}
\input{tools}


\section{Custom Framework}
In this section we describe the main functionalities of our
framework. We first give an overview of our proposed solution, and
then we detail the concrete extensions we have performed on the PIL
language. The description of the PIL Front-End and the target
Back-Ends are also covered in this section.
\label{customframework}
\input{customframework}


\section{Discussion}
\label{discussion}
\input{discussion}

\section{Use Case: Schnorr's Protocol}
\label{schnorr}
\input{schnorr}


%\section{Use Case: Direct Anonymous Attestation}
%\input{daa}
%\label{daa}

\section{Conclusion and Future Work}
\label{conclusion}
\input{conclusion}


\bibliographystyle{splncs}
\bibliography{main}

\appendix
\section{ANTLR: A Parser Generator}
\input{antlr}
\label{sec:antlr}
%\input{appendix/pk_details}

\end{document}

%%% Local Variables:
%%% TeX-PDF-mode: t
%%% TeX-master: t
%%% End:
