\documentclass[master=eelt, english]{kulemt}

\setup{
title={Implementation and Evaluation of Zero-Knowledge Proofs of Knowledge},
author={Boran Car},
promotor={Prof.\,dr.\,ir. Bart Preneel \and Prof.\,dr.\,ir. Ingrid Verbauwhede},
assessor={Prof.\,dr.\,ir. Claudia Diaz \and Prof.\,dr.\,ir. Francky Catthoor},
assistant={Ir. Josep Balasch \and Ir. Alfredo Rial},
}

\usepackage[pdfusetitle,colorlinks,plainpages=false]{hyperref}
\usepackage{biblatex}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{tikz-uml}
\usepackage[caption=false]{subfig}
\usepackage{tabularx}
\usepackage{color}
\usepackage{units}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{thm}{Theorem}
\newtheorem{corr}{Corollary}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstdefinelanguage{PIL}{
morekeywords={ExecutionOrder, Common, Prover, Verifier, Def, Void, Prime, Z, Int},
otherkeywords={Zmod+, Zmod*},
morecomment=[l]{//},
morecomment=[s]{/*}{*/}
}

\lstdefinelanguage{PSL}{
  morekeywords={Prime, Public, ProverPrivate,
    KnowledgeError, SZKParameter, ProtocolComposition, Homomorphism,
    ChallengeLength, SigmaPhi, Relation, Declarations, Inputs, Properties},
  otherkeywords={Zmod+, Zmod*},
  morecomment=[l]{//}, morecomment=[s]{/*}{*/}
}

\lstdefinelanguage{grammar}{
  keywords={},
  otherkeywords={}
}

\lstdefinelanguage{GEZEL}{
  morekeywords={sfg, dp, always, reg, syg, system, ns, tc, out, in, use, sig},
  otherkeywords={$display, $hex, $dec, $bin, $cycle},
  morecomment=[l]{//}
}

\lstset{
language=PIL,
basicstyle=\footnotesize,
commentstyle=\color{dkgreen},
keywordstyle=\bfseries,
frame=single,
breaklines=true,
breakatwhitespace=false
}

\usetikzlibrary{backgrounds,calc,shapes,shapes.geometric,shadows,arrows,chains,trees,fit,positioning}

\tikzstyle{language}=[rectangle,draw=black,thin,inner sep=3pt, align=center]

\tikzstyle{compiler}=[rectangle,rounded corners,draw=black,thin,inner
sep=3pt, align=center]

\tikzstyle{added}=[fill=gray]

\addbibresource{references}

\begin{document}

\begin{preface}

\end{preface}

\tableofcontents*

\begin{abstract}
  Zero-Knowledge Proofs of Knowledge are protocols that allow one to
  prove possession or knowing of a secret without actually revealing
  it. Goals of current Zero-Knowledge Proofs of Knowledge compiler
  frameworks are to make the implementation of these protocols easy
  for those outside the field of cryptography while at the same time
  making it easy for those inside the field to prototype such
  protocols. We acknowledge that such frameworks have made it easy to
  develop the protocols but we note that they have not made it easy or
  general to those implementing the protocols on the end devices. Such
  frameworks either target C/C++ code generation or an interpreted
  language. They also use high-level multi-precision arithmetic
  libraries which are too resource demanding for small embedded
  devices. These limitations as well as hidden pitfalls of C/C++ have
  motivated us to extend one of those frameworks and build our own
  framework from it. This framework is targeted specifically at HW-SW
  co-design exploration while not losing any generality of the
  framework it extends. We hope that this will ease the work for those
  outside the field and allow for a more widespread usage of these
  protocols.
\end{abstract}

\listoffiguresandtables

%\listoftables

\mainmatter

\input{introduction}

\input{preliminaries}

\input{existing}

\input{custom}

\input{usecase}

\input{conclusion}

\appendixpage*

\appendix

\input{coproc_report}

%\lstinputlisting{pil.g}

%\lstinputlisting{codegen.g}

\backmatter

\printbibliography

\end{document}

%%% Local Variables:
%%% TeX-PDF-mode: t
%%% TeX-master: t
%%% End:
